Functional timeseries reveal no systemic shifts in vertebrate communities

### 2021-06-21 paper/manuscript.Rmd
Add Carl comments

---
title: "Functional timeseries reveal no systemic shifts for vertebrate communities"
author: 
  - name: "Kari E A Norman"
    affiliation: ucb
  - name: "Carl Boettiger"
    affiliation: ucb, ropensci
address:
  - code: ucb
    address: "Dept of Environmental Science, Policy, and Management, University of California Berkeley, Berkeley CA 94720-3114, USA"
  - code: ropensci
    address: "The rOpenSci Project, University of California Berkeley, Berkeley CA 94720-3114, USA" 
abstract: |

@@

*Methodological Limitations*

This study represents best practices in biodiversity synthesis and functional diversity analysis, however methodological limitations may still be obscuring potential systematic shifts in functional composition.
First, as previously acknowledged, the Biotime database is not a representative sample of the world's biodiversity or areas of greatest threat [@gonzalez2016; @vellend2017], and the subset of data in this study exhibits a strong Northern Hemisphere bias.
While both of these challenges are pervasive in Ecology generally and could have potentially insidious implications for synthesis work, we can only use the data we have.
We do however emphasize that our study included many timeseries that experienced a significant species loss (\~ 52%), indicating that this is not simply a synthesis of pristine systems.

@@

*Implications for Ecological Process*

@@

@@

>>>>

---
title: "Functional timeseries reveal no systemic shifts in vertebrate communities"
author: 
  - name: "Kari E A Norman"
    affiliation: ucb
  - name: "Carl Boettiger"
    affiliation: ucb, ropensci
  - name: "Perry de Valpine"
    affiliation: ucb
address:
  - code: ucb
    address: "Dept of Environmental Science, Policy, and Management, University of California Berkeley, Berkeley CA 94720-3114, USA"
  - code: ropensci
    address: "The rOpenSci Project, University of California Berkeley, Berkeley CA 94720-3114, USA" 
abstract: |

@@

[@osullivan2021]

<!-- Link turnover paragraph to background turnover rate, maintenence of species diversity (i.e. turnover is happening! what does that mean for speices div? maintained. for functional diversity? maintained.) -->

*Implications for Future Work*

Here's the prediction of the field, here is the data of the field, given that the data doesn't match the prediction it's on the field to resolve these issues.

Why do we find no significant trend in the global data despite extensive predictions to the contrary from both theory and local studies?
First, there are a variety of data limitations to consider:

This study represents best practices in biodiversity synthesis and functional diversity analysis, however methodological limitations may still be obscuring potential systematic shifts in functional composition.
First, the Biotime database is not a representative sample of the world's biodiversity or areas of greatest threat [@gonzalez2016; @vellend2017], and the subset of data in this study exhibits a strong Northern Hemisphere bias.
While both of these challenges are pervasive in Ecology generally and could have potentially insidious implications for synthesis work, we can only use the data we have.
We do however emphasize that our study included many timeseries that experienced a significant species loss (\~ 52%), indicating that this is not simply a synthesis of pristine systems.

You'd imagine the pattern would still be visible even with North American birds...

@@

-   one opening sentence that links back to the null result

Even with the right data and metrics, you may not see a result!
Due to the wide variety of mechanisms at play we amy not expect to see a global trend even when the local trends are significant.

@@

We can not understand all functional loss in terms of species loss

-   lean in to four points in the abstraction

    -   raises questions for future data collection and theoretical development

-   field that's full of idle speculation

-   find papers that make claims we're "testing"

@@

They are limitations of the field not our limitations

-   acknowledge those things without taking the blame

-   lean into it as a conclusion

-   we found a null result and its your fault

-   fixing x, y, z could resolve the null result, but there are good reasons it might not

-   talk about balance of local trends as the fourth reason

There are reasons we haven't got trait data, hoped there would be an easy way out, but these results show that there isn't

Is there a null result for some deep theoretical reason or

we could've been lucky that we would see the pattern with crappy data, but it isn't

-   the thing we thought would be there isn't there, so what can we do to solve it

clever way to do temporal coverage - maps by decade?

-   number of data points through time (bar chart colored by taxa)

figure 1.
illustrating functional diversity metrics

figure 2.
the processing workflow
